### 2020-03-04 Common/chapters/paper_reproduction.tex
Johan and Julien: all names for each section added

\chapter{Research paper experiment reproduction}
\label{ch:paper_reproduction}
This chapter is based on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI" from Song et al.~\cite{07}, shortly presented in chapter \ref{ch:literature_review}. It aims at reproducing the experiment of the paper in order to acquire the medical, theoretical and technical background before proposing a transfer learning method as a way to improve the classification using other body parts (see Chapter \ref{ch:transfer_learning}).\\
Song et al. \cite{07} proposed a deep convolutional neural network (DCNN) method to detect prostate cancer based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset. This dataset is one of the biggest dataset available for prostate cancer classification. The two different output classes of the latter are benign lesion (class 0) and malignant lesion (class 1). The dataset contains a total of 204 patients for the training (labels are known) and 208 patients for the challenge (labels are unknown). This paper explains all the steps to follow with the goal of building a computer-aided diagnosis system for prostate cancer detection. Moreover, it also provides results (AUC) about the algorithm performance on a test set they built themselves (which is not the same as the challenge test set). In fact, they split the PROSTATEx training set into a training set, a validation set and the mentioned test set. These results can be used as a benchmark to compare with the results of the reproduction of the experiment. However, the article gives no information about the results their model achieved on the official PROSTATEx challenge test set (208 patients without labels). Therefore, the reproduction of the experiment will fill that gap by taking part in actual challenge.

@@

\section{Process overview}
Schematically speaking, the entire experiment process can be represented as shown on Figure \ref{fig:paper_reproduction_process}. The first part of the experiment is about the reproduction of the paper experiment, whereas the second one is devoted to the participation in the SPIE-AAPM-NCI PRO\- STATEx challenge since the authors of the article did not take part in it.\\
The former makes use of the PROSTATEx training set. This dataset is only composed of samples whose clinical significance is provided (labeled data). The first step of the reproduction is the data processing. In fact, many processing steps are involved to transform the original DICOM files into NumPy arrays which can be fed to the neural network. Data processing includes lesion cropping, normalization, MRI images stacking, data augmentation, etc. (see Section \ref{sec:prostatex_data_processing} for further details).\\
Once the processing part is completed, the labeled data are split into a training set, a validation set and a test set using respectively 80\%, 10\% and 10\% of the available data for each subset. The training set samples are then fed to the neural network in batches. 
%, which updates its weight consequently. 
At the end of each epoch, the current model is tested on the validation set and the resulting metrics are plotted on Tensorboard (see \ref{paper_tensorboard}). The validation set is independent of the training set and has never been seen by the model, which gives an indication of how well the latter generalizes. Then, if the current model reached a higher AUC and a higher accuracy than the previous best model, it becomes the new best model and is saved. Furthermore, if the validation AUC is decreasing for a defined number of epochs, the learning rate is reduced to prevent overfitting and avoid a  progressive decrease in validation performance. This technique contributes to maintaining a good generalization level and optimizes the training with respect to the AUC. As soon as the training phase reaches the defined number of epochs, the model is tested on the test set with the authors' method called "enhanced prediction" to compute the AUC. The latter method consists in predicting an output for each augmented image before averaging the predictions for each lesion of each patient.

@@

\subsection{PROSTATEx: Data processing}
\label{sec:prostatex_data_processing}
This section describes all processing operations performed on the raw data until it passes through the neural network. All steps mentioned below are also valid for the second part of the experiment about the challenge (Section \ref{sec:classification_challenge}).

\subsubsection{Dataset description}
\label{sec:prostatex_dataset_description}
The SPIE-AAPM-NCI PROSTATEx Challenge dataset is publicly available on the Cancer Imaging Archive website~\cite{33, 34, 35}. This dataset is composed of multiparametric MRIs for a total of 204 training patients and 208 challenge patients. Images were taken under the sagittal, transverse and coronal planes. These MRIs are multiparametric because the same exam was performed using different radiological settings called "sequences". Each sequence allows to see human tissues in a different way on the resulting images. The sequences available within the scope of PROSTATEx are T2 weighted images (T2W), Diffusion weighted images (DWI), Apparent diffusion coefficient (ADC), Dynamic contrast enhanced (DCE), PD (Proton Density) and K-trans.

@@

\subsubsection{Methodology}
The methodology described in this section is the closest replication possible of the authors' data processing methodology. Most of it were reproductible in a similar way, apart from the manual lesion contouring which was performed by a qualified radiologist in the authors' case. Furthermore, the authors did not mention any participation in the official PROSTATEx challenge. In order to get an unbiased evaluation of the performance of our model, these images were also processed as explained below in order to take part in the challenge, but were not augmented. Nevertheless, the authors created their own test set by splitting the PROSTATEx training images into a training set ($80\%$), a validation set ($10\%$) and a test set ($10\%$). They mentioned that the test set was augmented $11$ times but gave no information about the training and validation sets. Since the training set is imbalanced (3 false for 1 true), the true class was augmented more times than the false class ($60$x) in order to create balanced training and validation sets.

\subsubsection{From DICOM to NumPy arrays}
\label{sec:DICOMtoNumPy}
Before anything else, T2W, DWI and ADC grayscale images were used, which means that DCE, PD and Ktrans images were left aside. Furthermore, DWI images regroup various parameters which create different subcategories. One of them is the so-called "b-value" which is "a factor that reflects the strength and timing of the gradients used to generate diffusion-weighted images. The higher the b-value, the stronger the diffusion effects" \cite{49}. The authors tested three different b-value configurations. This led to the conclusion that using DWI images with the highest b-value increased the performance of the model. For that matter, DWI images with the highest b-value only were used. \\
Moreover, only images showing the prostate under the transverse plane were used. The reason for this is that the tumors are much more visible under this perspective. 
The first step consisted in converting DICOM files. Algorithm \ref{alg:PROSTATEx_preprocessing} describes the steps involved in converting PROSTATEx's DICOM files to NumPy arrays. The right slices were found thanks to the two CSV files. Important information such as the patient ID, the sequence, the lesion location and the voxel spacing were used as file output names, which allows to use these files independently for the next steps (i.e. without relying on the CSV files).

@@

\subsubsection{From NumPy arrays to augmented stacked images}
\label{sec:numpyToAugmentedStacked}
This step can be split into two substeps: aligning and stacking images before augmenting them and splitting them across a training set, a validation set and a test set.

@@

\subsubsection{From NumPy arrays to augmented non-stacked images}
Another way of processing and using the data was tested. Instead of stacking the images of various sequences coming from the same exam, each image was treated independently. In other words, each of them was resized as explained previously, then cropped, augmented and exported separately. At the end, each image was fed as separate input to the neural network.

@@

\subsection{Data processing verification}
\label{sec:dataProcessingVerification}

\subsubsection{Cropping verification using red dots}
In order to check the correctness of the augmentation process, images with a red dot at the lesion coordinates were generated. First of all, a red dot was placed on the full image thanks to the "i,j,k" attribute in the CSV file. This full image was exported. Then, images were augmented in the exact same manner as described in section \ref{sec:numpyToAugmentedStacked}. The expected result satisfies the following properties:
\begin{itemize}
	\item The red dot is localized in the center of the cropped and augmented image.
	\item The three cropped and augmented images within a stacked image contain the exact same tissue. 
	\item The three cropped and augmented images within a stacked image are not rotated/shifted/flipped differently. 
	\item A lesion is visible to the naked eye. 
\end{itemize}

@@

\subsection{Training the neural network}
Training the neural network is the most important step of the experiment after the data processing. This section first describes the architecture of the model, then presents the script used for the training and finally provides details about the experimental setup such as the training procedure, the hyperparameters of the model and the configuration of the machine on which it was trained.

\subsubsection{Architecture}
The model architecture used in the paper is a modified version of the VGG network from the Oxford's Visual Geometry Group (VGG). This model was initially designed as part of the Large Scale Visual Recognition Challenge 2014 that used the ImageNet dataset composed of 14 millions images belonging to 1000 classes. Figure \ref{fig:paper_model} illustrates its structure and its corresponding implementation in Python with the PyTorch framework.\\
It is first composed of three convolution-dropout-max-pooling blocks followed by three fully-connected-dropout blocks. Each convolutional box (in blue) in the figure represents in reality three layers: the convolutional layer, the batch normalization layer and the exponential linear unit (ELU) activation function. The same principle applies to the fully connected layer box (in orange) that is divided into a fully connected layer followed by the exponential linear unit. The last fully connected box (in purple) has the same structure except that the exponential linear unit is replaced by a softmax function for classification.\\
In comparison to the original VGG, this model keeps the small filter size of 3x3, also doubles the number of filters after each convolution-dropout-max-pooling block and has a stride of 1 for all convolutions. It differs from the traditional VGG in that it makes use of a smaller number of layers since the task is simpler than the original one. Moreover, it uses exponential linear units instead of rectified linear units as activation functions and adds dropout and batch normalization layers. Also, it uses 1x1 convolutions.

@@

\subsubsection{Script options}
Since training a neural network requires datasets, hyperparameters and many more configuration choices, the creation of a generic script that accepts multiple options is necessary. This requirement was solved thanks to the Python module called "argparse". The latter automatically generates help/usage messages and displays errors when the arguments typed in by the user are invalid. Table \ref{fig:paper_reproduction_options} shows all options accepted by the script, provides information about what they concretely mean and indicates their respective types.

@@

\subsubsection{Tensorboard}
Tensorboard's official website \cite{39} claims that "Tensorboard provides the visualization and tooling needed for machine learning experimentation:
\begin{itemize}
\item Tracking and visualizing metrics.
\item Visualizing the model graph (ops and layers).
\item Viewing histograms of weights, biases, or other tensors as they change over time.
\item Projecting embeddings to a lower dimensional space.
\item Displaying images, text, and audio data"\cite{39}.
\end{itemize}
In this experiment, Tensorboard is mainly used to plot Matplotlib figures of the model performance. In fact, during the training the following metrics are computed: loss, accuracy, precision, recall, F1-score, specificity and AUC. These metrics are computed separately on the training and on the validation sets at the end of each epoch. They are then stored and plotted on the same figure at the end of the process thanks to Tensorboard. In addition to this, written reports regarding which model was the best and the results it achieved are also added to the dashboard.
\label{paper_tensorboard}

\subsubsection{Model roulette}
As stated by the authors, "the training process is sensitive to the parameter initialization" \cite{07}. Experience showed us that the exact same hyperparameters with two different initializations could give opposite results. Therefore, in order not to train the model vainly, a script which generates a given number of initializations was created. Each untrained model was then tested on the validation set since the best model during training is saved according to its performance on the validation set. The model whose score was the highest on a given metric (the AUC in this experiment) was saved and used as base model for the experiments. It can then be used as initialization model using the script option \textit{--modeltoload <path>}.

\subsubsection{Experimental setup}
Keeping the exact same hyperparameters as the original paper led the experiment to disappointing results. Consequently, new hyperparameters had to be found. A grid search approach driven by the AUC obtained on the test set (the one built for the experiment, not the one of the challenge) was chosen. In practical terms, this consisted in trying multiple combinations of hyperparameters and keeping the one that produced the best AUC on this test set. At the end of this process, it appeared that the best model was obtained when the data was organized in batches of 32 samples, with an initial learning rate of $1e-7$, a learning rate scheduler on plateau using a factor of 100 with a patience of 2 (the learning rate is divided by 100 when the validation AUC decreases for two consecutive epochs) and a dropout rate of 0.2. Note that the model was initialized picking the best one among 200 models generated from a roulette maximizing the AUC. This model was implemented with PyTorch using the Adam optimizer to update its weights and was trained on an Nvidia GeForce GTX Titan X graphics card during approximately 3 hours.\\

\subsection{Training verification}
In deep learning, two main problems can occur during training. The first one is called the "vanishing gradient" problem and refers to the fact that it is possible for the loss function to compute extremely small gradients, near zero. Consequently, the weight update is also extremely small, which makes the neural network hard or impossible to train. The second problem, the "exploding gradient" is the opposite. In this case, large gradients are computed, which leads to huge weight updates that can even reach "NaN" values. This makes the model unstable and unable to learn from training data.  \\
As a result, it is important to analyse the gradient propagation across the network to ensure that it is not facing such problems. To achieve this goal, a visualization tool to display the gradients of the entire network was implemented. This implementation is based on the code of RoshanRane \cite{47}.  This allows to visually notice the gradient flow at a glance and to see its evolution during all epochs (one visualization per batch is generated).

@@

\subsection{Results}
\label{sec:paper_reproduction_results}
Figure \ref{fig:paper_reprodution_results} shows the different metrics registered on the training and validation sets during the training phase. The best model was chosen at epoch 21 with an accuracy of $0.7525$ and an AUC of $0.765$ on the validation set (recall: the condition for a model to be chosen as best is that the current AUC and accuracy should be higher than the AUC and accuracy of the previous best model).\\
This best model was evaluated on the test set built for the experiment and achieved an AUC of $0.75$ using their "enhanced prediction" technique (see figure \ref{fig:paper_reprodution_auc_on_our_test_set}).\\
Finally, the model was also tested on the official SPIE-AAPM-NCI Prostate MR Classification Challenge test set and achieved an AUC of 0.71 (see figure \ref{fig:paper_reprodution_results_challenge_1}). This result will be used as benchmark in comparison with the model trained on the whole dataset in the second part of the experiment.

@@

\subsection{Discussion}
Thanks to the curves of Section \ref{sec:paper_reproduction_results}, a lot of information regarding the model performance can be inferred. In fact, "reviewing learning curves of models during training can be used to diagnose problems with learning, such as an underfit or overfit model, as well as whether the training and validation datasets are suitably representative"\cite{40}.\\
First of all, when observing the loss plot, what is noticeable is that the training and validation losses have the same behavior during the 50 epochs. This is a clear sign of good fit of the model on the data. All the features learned by the model on the training data are also applicable to the validation set. This situation is the optimal one. Regarding the slope of the curve, a big change occurs at epoch 17. This moment corresponds to the one where the learning rate is reduced by a factor of $100$ thanks to the \textit{reducelronplateau} function. In fact, at this moment, the validation AUC (the training is optimized according to this metric) was at its optimum with this learning rate. If this technique had not been used, the training loss would have continued to go down whereas the validation loss would have gone up immediately, which would have lead to overfitting. Here, thanks to the learning rate scheduler, this scenario is avoided and the validation and training losses stay stable from this point onward. The same behaviour is clearly noticeable for all other metrics.\\
Regarding the overall performance of the best model chosen at epoch 21, the latter generalizes well. This is particularly visible on the accuracy and AUC plots where the validation metrics are higher or equal to the training ones. The validation accuracy indicates that the model made the right class decision in approximately $75$\% of the time. Furthermore, since the model achieved a validation AUC of $0.76$, it shows that it had a good distinction capacity. In fact, the highest the AUC is, the highest is the ability of the model to predict benign lesions as benign and malignant lesions as malignant. Samuel G. et al. \cite{42} claim that a less-experienced radiologist can reach an AUC of $0.81$ and an expert an AUC of $0.91$. Consequently, this model is almost at the level of a radiologist, which is very promising. It is also important to keep in mind that the model was trained using only 164 patients (80\% of the 204 training patients), which suggests that even better results can be reached with more data.\\
The precision plot, which is used to measure the proportion of malignant lesions correctly classified among all lesions classified as malignant lesions, confirms the hypothesis that the model has a good distinction capacity with a score of approximately $0.76$ on the validation set.\\
Another extremely important aspect of a computer-aider diagnosis system is to avoid the number of false negatives as much as possible. In fact, the consequences of classifying a malignant lesion as benign are far worse than the opposite. This concept is measured by the recall, which quantifies the proportion of malignant lesions correctly classified among all malignant lesion. The model has a recall of around $0.74$ on the validation set.\\
The two last metrics (precision and recall) are both considered in the f1-score that is equal to $0.76$. \\
Finally, the specificity indicates the proportion of correctly classified benign lesions among all benign lesions and is equal to approximately $0.76$.\\
Globally, this model is balanced as no metric is exploding at the expense of another. This stability constitutes a serious strength regarding its generalization ability.

@@

\section{SPIE-AAPM-NCI Prostate MR Classification Challenge}
\label{sec:classification_challenge}
This section aims at maximizing the result on the SPIE-AAPM-NCI Prostate MR classification Challenge test set by using the entire dataset as training set, instead of using 80\% of it, as it was done in the previous section. First, the changes needed to reach this new goal are detailed, then the raw results are presented and the latter are finally interpreted.

>>>>

\chapter{Research paper experiment reproduction}
\label{ch:paper_reproduction}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}This chapter is based on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI" from Song et al.~\cite{07}, shortly presented in chapter \ref{ch:literature_review}. It aims at reproducing the experiment of the paper in order to acquire the medical, theoretical and technical background before proposing a transfer learning method as a way to improve the classification using other body parts (see Chapter \ref{ch:transfer_learning}).\\
Song et al. \cite{07} proposed a deep convolutional neural network (DCNN) method to detect prostate cancer based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset. This dataset is one of the biggest dataset available for prostate cancer classification. The two different output classes of the latter are benign lesion (class 0) and malignant lesion (class 1). The dataset contains a total of 204 patients for the training (labels are known) and 208 patients for the challenge (labels are unknown). This paper explains all the steps to follow with the goal of building a computer-aided diagnosis system for prostate cancer detection. Moreover, it also provides results (AUC) about the algorithm performance on a test set they built themselves (which is not the same as the challenge test set). In fact, they split the PROSTATEx training set into a training set, a validation set and the mentioned test set. These results can be used as a benchmark to compare with the results of the reproduction of the experiment. However, the article gives no information about the results their model achieved on the official PROSTATEx challenge test set (208 patients without labels). Therefore, the reproduction of the experiment will fill that gap by taking part in actual challenge.

@@

\section{Process overview}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Schematically speaking, the entire experiment process can be represented as shown on Figure \ref{fig:paper_reproduction_process}. The first part of the experiment is about the reproduction of the paper experiment, whereas the second one is devoted to the participation in the SPIE-AAPM-NCI PRO\- STATEx challenge since the authors of the article did not take part in it.\\
The former makes use of the PROSTATEx training set. This dataset is only composed of samples whose clinical significance is provided (labeled data). The first step of the reproduction is the data processing. In fact, many processing steps are involved to transform the original DICOM files into NumPy arrays which can be fed to the neural network. Data processing includes lesion cropping, normalization, MRI images stacking, data augmentation, etc. (see Section \ref{sec:prostatex_data_processing} for further details).\\
Once the processing part is completed, the labeled data are split into a training set, a validation set and a test set using respectively 80\%, 10\% and 10\% of the available data for each subset. The training set samples are then fed to the neural network in batches. 
%, which updates its weight consequently. 
At the end of each epoch, the current model is tested on the validation set and the resulting metrics are plotted on Tensorboard (see \ref{paper_tensorboard}). The validation set is independent of the training set and has never been seen by the model, which gives an indication of how well the latter generalizes. Then, if the current model reached a higher AUC and a higher accuracy than the previous best model, it becomes the new best model and is saved. Furthermore, if the validation AUC is decreasing for a defined number of epochs, the learning rate is reduced to prevent overfitting and avoid a  progressive decrease in validation performance. This technique contributes to maintaining a good generalization level and optimizes the training with respect to the AUC. As soon as the training phase reaches the defined number of epochs, the model is tested on the test set with the authors' method called "enhanced prediction" to compute the AUC. The latter method consists in predicting an output for each augmented image before averaging the predictions for each lesion of each patient.

@@

\subsection{PROSTATEx: Data processing}
\label{sec:prostatex_data_processing}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}This section describes all processing operations performed on the raw data until it passes through the neural network. All steps mentioned below are also valid for the second part of the experiment about the challenge (Section \ref{sec:classification_challenge}).

\subsubsection{Dataset description}
\label{sec:prostatex_dataset_description}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}The SPIE-AAPM-NCI PROSTATEx Challenge dataset is publicly available on the Cancer Imaging Archive website~\cite{33, 34, 35}. This dataset is composed of multiparametric MRIs for a total of 204 training patients and 208 challenge patients. Images were taken under the sagittal, transverse and coronal planes. These MRIs are multiparametric because the same exam was performed using different radiological settings called "sequences". Each sequence allows to see human tissues in a different way on the resulting images. The sequences available within the scope of PROSTATEx are T2 weighted images (T2W), Diffusion weighted images (DWI), Apparent diffusion coefficient (ADC), Dynamic contrast enhanced (DCE), PD (Proton Density) and K-trans.

@@

\subsubsection{Methodology}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}
The methodology described in this section is the closest replication possible of the authors' data processing methodology. Most of it were reproductible in a similar way, apart from the manual lesion contouring which was performed by a qualified radiologist in the authors' case. Furthermore, the authors did not mention any participation in the official PROSTATEx challenge. In order to get an unbiased evaluation of the performance of our model, these images were also processed as explained below in order to take part in the challenge, but were not augmented. Nevertheless, the authors created their own test set by splitting the PROSTATEx training images into a training set ($80\%$), a validation set ($10\%$) and a test set ($10\%$). They mentioned that the test set was augmented $11$ times but gave no information about the training and validation sets. Since the training set is imbalanced (3 false for 1 true), the true class was augmented more times than the false class ($60$x) in order to create balanced training and validation sets.

\subsubsection{From DICOM to NumPy arrays}
\label{sec:DICOMtoNumPy}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Before anything else, T2W, DWI and ADC grayscale images were used, which means that DCE, PD and Ktrans images were left aside. Furthermore, DWI images regroup various parameters which create different subcategories. One of them is the so-called "b-value" which is "a factor that reflects the strength and timing of the gradients used to generate diffusion-weighted images. The higher the b-value, the stronger the diffusion effects" \cite{49}. The authors tested three different b-value configurations. This led to the conclusion that using DWI images with the highest b-value increased the performance of the model. For that matter, DWI images with the highest b-value only were used. \\
Moreover, only images showing the prostate under the transverse plane were used. The reason for this is that the tumors are much more visible under this perspective. 
The first step consisted in converting DICOM files. Algorithm \ref{alg:PROSTATEx_preprocessing} describes the steps involved in converting PROSTATEx's DICOM files to NumPy arrays. The right slices were found thanks to the two CSV files. Important information such as the patient ID, the sequence, the lesion location and the voxel spacing were used as file output names, which allows to use these files independently for the next steps (i.e. without relying on the CSV files).

@@

\subsubsection{From NumPy arrays to augmented stacked images}
\label{sec:numpyToAugmentedStacked}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}This step can be split into two substeps: aligning and stacking images before augmenting them and splitting them across a training set, a validation set and a test set.

@@

\subsubsection{From NumPy arrays to augmented non-stacked images}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Another way of processing and using the data was tested. Instead of stacking the images of various sequences coming from the same exam, each image was treated independently. In other words, each of them was resized as explained previously, then cropped, augmented and exported separately. At the end, each image was fed as separate input to the neural network.

@@

\subsection{Data processing verification}
\label{sec:dataProcessingVerification}
\subsubsection{Cropping verification using red dots}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}In order to check the correctness of the augmentation process, images with a red dot at the lesion coordinates were generated. First of all, a red dot was placed on the full image thanks to the "i,j,k" attribute in the CSV file. This full image was exported. Then, images were augmented in the exact same manner as described in section \ref{sec:numpyToAugmentedStacked}. The expected result satisfies the following properties:
\begin{itemize}
	\item The red dot is localized in the center of the cropped and augmented image.
	\item The three cropped and augmented images within a stacked image contain the exact same tissue. 
	\item The three cropped and augmented images within a stacked image are not rotated/shifted/flipped differently. 
	\item A lesion is visible to the naked eye. 
\end{itemize}

@@

\subsection{Training the neural network}

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Training the neural network is the most important step of the experiment after the data processing. This section first describes the architecture of the model, then presents the script used for the training and finally provides details about the experimental setup such as the training procedure, the hyperparameters of the model and the configuration of the machine on which it was trained.

\subsubsection{Architecture}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}The model architecture used in the paper is a modified version of the VGG network from the Oxford's Visual Geometry Group (VGG). This model was initially designed as part of the Large Scale Visual Recognition Challenge 2014 that used the ImageNet dataset composed of 14 millions images belonging to 1000 classes. Figure \ref{fig:paper_model} illustrates its structure and its corresponding implementation in Python with the PyTorch framework.\\
It is first composed of three convolution-dropout-max-pooling blocks followed by three fully-connected-dropout blocks. Each convolutional box (in blue) in the figure represents in reality three layers: the convolutional layer, the batch normalization layer and the exponential linear unit (ELU) activation function. The same principle applies to the fully connected layer box (in orange) that is divided into a fully connected layer followed by the exponential linear unit. The last fully connected box (in purple) has the same structure except that the exponential linear unit is replaced by a softmax function for classification.\\
In comparison to the original VGG, this model keeps the small filter size of 3x3, also doubles the number of filters after each convolution-dropout-max-pooling block and has a stride of 1 for all convolutions. It differs from the traditional VGG in that it makes use of a smaller number of layers since the task is simpler than the original one. Moreover, it uses exponential linear units instead of rectified linear units as activation functions and adds dropout and batch normalization layers. Also, it uses 1x1 convolutions.

@@

\subsubsection{Script options}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}
Since training a neural network requires datasets, hyperparameters and many more configuration choices, the creation of a generic script that accepts multiple options is necessary. This requirement was solved thanks to the Python module called "argparse". The latter automatically generates help/usage messages and displays errors when the arguments typed in by the user are invalid. Table \ref{fig:paper_reproduction_options} shows all options accepted by the script, provides information about what they concretely mean and indicates their respective types.

@@

\subsubsection{Tensorboard}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Tensorboard's official website \cite{39} claims that "Tensorboard provides the visualization and tooling needed for machine learning experimentation:
\begin{itemize}
\item Tracking and visualizing metrics.
\item Visualizing the model graph (ops and layers).
\item Viewing histograms of weights, biases, or other tensors as they change over time.
\item Projecting embeddings to a lower dimensional space.
\item Displaying images, text, and audio data"\cite{39}.
\end{itemize}
In this experiment, Tensorboard is mainly used to plot Matplotlib figures of the model performance. In fact, during the training the following metrics are computed: loss, accuracy, precision, recall, F1-score, specificity and AUC. These metrics are computed separately on the training and on the validation sets at the end of each epoch. They are then stored and plotted on the same figure at the end of the process thanks to Tensorboard. In addition to this, written reports regarding which model was the best and the results it achieved are also added to the dashboard.
\label{paper_tensorboard}

\subsubsection{Model roulette}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}As stated by the authors, "the training process is sensitive to the parameter initialization" \cite{07}. Experience showed us that the exact same hyperparameters with two different initializations could give opposite results. Therefore, in order not to train the model vainly, a script which generates a given number of initializations was created. Each untrained model was then tested on the validation set since the best model during training is saved according to its performance on the validation set. The model whose score was the highest on a given metric (the AUC in this experiment) was saved and used as base model for the experiments. It can then be used as initialization model using the script option \textit{--modeltoload <path>}.

\subsubsection{Experimental setup}

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}
Keeping the exact same hyperparameters as the original paper led the experiment to disappointing results. Consequently, new hyperparameters had to be found. A grid search approach driven by the AUC obtained on the test set (the one built for the experiment, not the one of the challenge) was chosen. In practical terms, this consisted in trying multiple combinations of hyperparameters and keeping the one that produced the best AUC on this test set. At the end of this process, it appeared that the best model was obtained when the data was organized in batches of 32 samples, with an initial learning rate of $1e-7$, a learning rate scheduler on plateau using a factor of 100 with a patience of 2 (the learning rate is divided by 100 when the validation AUC decreases for two consecutive epochs) and a dropout rate of 0.2. Note that the model was initialized picking the best one among 200 models generated from a roulette maximizing the AUC. This model was implemented with PyTorch using the Adam optimizer to update its weights and was trained on an Nvidia GeForce GTX Titan X graphics card during approximately 3 hours.\\

\subsection{Training verification}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}In deep learning, two main problems can occur during training. The first one is called the "vanishing gradient" problem and refers to the fact that it is possible for the loss function to compute extremely small gradients, near zero. Consequently, the weight update is also extremely small, which makes the neural network hard or impossible to train. The second problem, the "exploding gradient" is the opposite. In this case, large gradients are computed, which leads to huge weight updates that can even reach "NaN" values. This makes the model unstable and unable to learn from training data.  \\
As a result, it is important to analyse the gradient propagation across the network to ensure that it is not facing such problems. To achieve this goal, a visualization tool to display the gradients of the entire network was implemented. This implementation is based on the code of RoshanRane \cite{47}.  This allows to visually notice the gradient flow at a glance and to see its evolution during all epochs (one visualization per batch is generated).

@@

\subsection{Results}
\label{sec:paper_reproduction_results}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Figure \ref{fig:paper_reprodution_results} shows the different metrics registered on the training and validation sets during the training phase. The best model was chosen at epoch 21 with an accuracy of $0.7525$ and an AUC of $0.765$ on the validation set (recall: the condition for a model to be chosen as best is that the current AUC and accuracy should be higher than the AUC and accuracy of the previous best model).\\
This best model was evaluated on the test set built for the experiment and achieved an AUC of $0.75$ using their "enhanced prediction" technique (see figure \ref{fig:paper_reprodution_auc_on_our_test_set}).\\
Finally, the model was also tested on the official SPIE-AAPM-NCI Prostate MR Classification Challenge test set and achieved an AUC of 0.71 (see figure \ref{fig:paper_reprodution_results_challenge_1}). This result will be used as benchmark in comparison with the model trained on the whole dataset in the second part of the experiment.

@@

\subsection{Discussion}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Thanks to the curves of Section \ref{sec:paper_reproduction_results}, a lot of information regarding the model performance can be inferred. In fact, "reviewing learning curves of models during training can be used to diagnose problems with learning, such as an underfit or overfit model, as well as whether the training and validation datasets are suitably representative"\cite{40}.\\
First of all, when observing the loss plot, what is noticeable is that the training and validation losses have the same behavior during the 50 epochs. This is a clear sign of good fit of the model on the data. All the features learned by the model on the training data are also applicable to the validation set. This situation is the optimal one. Regarding the slope of the curve, a big change occurs at epoch 17. This moment corresponds to the one where the learning rate is reduced by a factor of $100$ thanks to the \textit{reducelronplateau} function. In fact, at this moment, the validation AUC (the training is optimized according to this metric) was at its optimum with this learning rate. If this technique had not been used, the training loss would have continued to go down whereas the validation loss would have gone up immediately, which would have lead to overfitting. Here, thanks to the learning rate scheduler, this scenario is avoided and the validation and training losses stay stable from this point onward. The same behaviour is clearly noticeable for all other metrics.\\
Regarding the overall performance of the best model chosen at epoch 21, the latter generalizes well. This is particularly visible on the accuracy and AUC plots where the validation metrics are higher or equal to the training ones. The validation accuracy indicates that the model made the right class decision in approximately $75$\% of the time. Furthermore, since the model achieved a validation AUC of $0.76$, it shows that it had a good distinction capacity. In fact, the highest the AUC is, the highest is the ability of the model to predict benign lesions as benign and malignant lesions as malignant. Samuel G. et al. \cite{42} claim that a less-experienced radiologist can reach an AUC of $0.81$ and an expert an AUC of $0.91$. Consequently, this model is almost at the level of a radiologist, which is very promising. It is also important to keep in mind that the model was trained using only 164 patients (80\% of the 204 training patients), which suggests that even better results can be reached with more data.\\
The precision plot, which is used to measure the proportion of malignant lesions correctly classified among all lesions classified as malignant lesions, confirms the hypothesis that the model has a good distinction capacity with a score of approximately $0.76$ on the validation set.\\
Another extremely important aspect of a computer-aider diagnosis system is to avoid the number of false negatives as much as possible. In fact, the consequences of classifying a malignant lesion as benign are far worse than the opposite. This concept is measured by the recall, which quantifies the proportion of malignant lesions correctly classified among all malignant lesion. The model has a recall of around $0.74$ on the validation set.\\
The two last metrics (precision and recall) are both considered in the f1-score that is equal to $0.76$. \\
Finally, the specificity indicates the proportion of correctly classified benign lesions among all benign lesions and is equal to approximately $0.76$.\\
Globally, this model is balanced as no metric is exploding at the expense of another. This stability constitutes a serious strength regarding its generalization ability.

@@

\section{SPIE-AAPM-NCI Prostate MR Classification Challenge}
\label{sec:classification_challenge}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}This section aims at maximizing the result on the SPIE-AAPM-NCI Prostate MR classification Challenge test set by using the entire dataset as training set, instead of using 80\% of it, as it was done in the previous section. First, the changes needed to reach this new goal are detailed, then the raw results are presented and the latter are finally interpreted.

\subsection{Training the neural network with the whole dataset}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}The first part of the experiment, in which the whole dataset was split into a training (80\%), a validation (10\%) and a test set (10\%), was extremely useful to find the right hyperparameters. It also gave the opportunity to check that the model and the processing of the data worked as expected with the goal of creating a state-of-the-art deep learning model for prostate lesions classification.\\
As a complement, this part of the experiment focuses on the challenge results in order to improve the $0.71$ in AUC achieved by the model trained with 80\% of all available data. In a general way, the more available data there is, the better the model performs. Consequently, the idea behind this section is to use the entire dataset as training set, discarding the validation and test sets from previous section. Apart from this new split, the data processing and the hyperparameters stay the same. The only difference in the training is about the learning rate scheduler. Since the validation set does not exist anymore, there is no possibility to use the \textit{reducelronplateau} function to prevent overfitting. However, this does not cause any problem thanks to the results of the previous section. In fact, it is clearly visible that the model would have started to overfit at epoch 17 if no learning rate correction had been applied. Hence, it is possible to conclude that, in this case, the best model should also be the one around epoch 17 (or a few epochs later). In order to ensure this, the models at epochs 15, 16, 17, 18, 19, 20, 21, 22, 23 and 30 were evaluated on the official challenge test set. For each of these models, a CSV file containing the predictions for each lesion of each patient was automatically generated and submitted to the challenge.
